Exact asymptotic expressions for local clustering in hyperbolic random graphs

### 2018-11-05 concentration.tex
fully merged compations of Nikolaos plus some small edits on section 2
Section 4 done (plus some small changes to Section 1.5)

\begin{proof}[Proof of Proposition \ref{prop:concentration_tilde_T_P_n}]
The technical steps of this proof use Lemma \ref{lem:joint_degree_distribution_shift} and Lemma \ref{lem:joint_degree_distribution_P_n}, which deal with the joint degree distribution. Let $\Ncal_{\Pcal,n}^c(p,p^\prime)$ denote the number of nodes not connected to both $p$ and $p^\prime$ an fix $0 < \eps < (2\alpha - 1 \wedge 1)$. To be able to apply both lemmas, we first show that we only need to consider the case where $\Exp{\left|\Ncal_{\Pcal,n}^c(p,p^\prime)\right|} \ge k_n^{\varepsilon}$ and $|x - x^\prime| > k_n^{1+\eps}$. To this end define 
\[
	\mathcal{E}_{\varepsilon}(k_n) = \left\{(p,p^\prime) \in \Kcal_C(k_n) \times \Kcal_C(k_n) 
	\, : \, \Exp{\left|\Ncal_{\Pcal,n}^c(p,p^\prime)\right|} \ge k_n^{\varepsilon} \text{ and } |x - x^\prime| > k_n^{1 + \varepsilon} \right\}
\]
and let $I_{a,b}^\ast$ denote the the part of $I_{a,b}$ where $p,p^\prime \in \Pcal_n \cap \mathcal{E}_\varepsilon(k_n)$. We first show that
\begin{eqnarray}\label{eq:I_ab_ast_main}
	\Exp{I_{a,b} - I_{a,b}^\ast} = \smallO{\Exp{T_{\Pcal,n}(k_n)}^2},
\end{eqnarray}
so that for the remainder of the proof we only need to consider $p, p^\prime \in \mathcal{E}_\varepsilon(k_n)$ and hence, we can apply Lemma \ref{lem:joint_degree_distribution_shift} and Lemma \ref{lem:joint_degree_distribution_P_n} to the joint degree distribution inside the integral.

@@

The proofs for the other two cases $I_{0,2}^\ast$ and $I_{0,2}^\ast$ follows similar arguments.

@@

\PvdH{Stopped here with general pass}

>>>>

\begin{proof}[Proof of Proposition \ref{prop:concentration_tilde_T_P_n}]

\PvdH{Include the case $k_n = \bigT{1}$.}

The technical steps of this proof use Lemma \ref{lem:joint_degree_distribution_shift} and Lemma \ref{lem:joint_degree_distribution_P_n}, which deal with the joint degree distribution. Let $\Ncal_{\Pcal,n}^c(p,p^\prime)$ denote the number of nodes not connected to both $p$ and $p^\prime$ an fix $0 < \eps < (2\alpha - 1 \wedge 1)$. To be able to apply both lemmas, we first show that we only need to consider the case where $\Exp{\left|\Ncal_{\Pcal,n}^c(p,p^\prime)\right|} \ge k_n^{\varepsilon}$ and $|x - x^\prime| > k_n^{1+\eps}$. To this end define 
\[
	\mathcal{E}_{\varepsilon}(k_n) = \left\{(p,p^\prime) \in \Kcal_C(k_n) \times \Kcal_C(k_n) 
	\, : \, \Exp{\left|\Ncal_{\Pcal,n}^c(p,p^\prime)\right|} \ge k_n^{\varepsilon} \text{ and } |x - x^\prime| > k_n^{1 + \varepsilon} \right\}
\]
and let $I_{a,b}^\ast$ denote the the part of $I_{a,b}$ where $p,p^\prime \in \Pcal_n \cap \mathcal{E}_\varepsilon(k_n)$. We first show that
\begin{eqnarray}\label{eq:I_ab_ast_main}
	\Exp{I_{a,b} - I_{a,b}^\ast} = \smallO{\Exp{T_{\Pcal,n}(k_n)}^2},
\end{eqnarray}
so that for the remainder of the proof we only need to consider $p, p^\prime \in \mathcal{E}_\varepsilon(k_n)$ and hence, we can apply Lemma \ref{lem:joint_degree_distribution_shift} and Lemma \ref{lem:joint_degree_distribution_P_n} to the joint degree distribution inside the integral.

@@

The proofs for the other two cases $I_{0,2}^\ast$ and $I_{0,2}^\ast$ follow using similar arguments.

@@

### 2018-11-05 introduction.tex
typo in intro/motivation and missing alphas in section 4.2
Section 4 done (plus some small changes to Section 1.5)

The aim of this paper is to study local clustering in hyperbolic random graphs. The first rigorous mathematical analysis of the local clustering coefficient was done by Gugelmann et al.~\cite{gugelmann2012random}. They show~\cite[Theorem 2.1]{gugelmann2012random} that the local clustering coefficient\footnote{Note that in~\cite{gugelmann2012random} this is called the global clustering coefficient. However, since this term is more often associated in the literature with the density of triangles compared to the number of paths of length 2, we use the term local clustering coefficient as is done, for instance, in~\cite{candellero2016clustering}.} is concentrated around its expectation which is shown to be $\bigT{1}$ as the number of vertices tends to infinity. In particular this results implies that the local clustering coefficient is asymptotically bounded away from zero. The convergence of this coefficient was not proven. In contrast, the global clustering coefficient was shown~\cite{candellero2016clustering} to converge in probability to some constant which can be explicitly stated as an integral expression, see Theorem 1.2 in \cite{candellero2016clustering}. The authors mention, however, that their analysis needs significant modification to be able to deal with convergence of the local clustering coefficient.

@@

The majority of the paper is dedicated to prove Theorem~\ref{thm:local_clustering_hyperbolic}, which is based on a sequence of subsequent results for the local clustering function in related random graph models and Theorem~\ref{thm:asymptotics_average_clustering_P}. The exact flow of the argument is explained in Section \ref{sec:proof_outline}. The proof for Theorem~\ref{thm:local_clustering_hyperbolic}, using these results, can be found in Section \ref{ssec:proof_main_results}. Here we also give the proof of Theorem~\ref{thm:clustering_coefficient_hyperbolic}. The proof of Theorem~\ref{thm:asymptotics_average_clustering_P} can be found in Section~\ref{sec:asymptotics_average_clustering_ast_P}.

@@

@@

\paragraph{Transition in the scaling of local clustering}
%Note that Theorem \ref{thm:local_clustering_hyperbolic} implies that for $\frac{1}{2} < \alpha < \frac{3}{4}$ the local clustering $c_{\H,n}(k)$ decays as $k^{-(4\alpha - 2)}$ while for $\alpha > \frac{3}{4}$ it is $k^{-1}$.

\paragraph{Universality of scaling transition}

\paragraph{Geometric inhomogeneous random graphs}

%\paragraph{Local clustering for growing versus fixed degrees}
%%The main difference between Theorem \ref{thm:fixed_local_clustering_hyperbolic} and Theorem \ref{thm:local_clustering_hyperbolic} is that for fixed degrees $k$ we have a small correction term that depends on the value $k$. 
%
%\MS{We note that the above statements / theorems are disjoint: in one case the degree is fixed corresponding to a constant sequence and in this case the error only depends on $n$, whereas in the other case, the degrees are assumed to be increasing, so in particular unbounded resp. tending to infinity and only the leading term in the asymptotic expansion w.r.t. $k_n \rightarrow \infty$ is given. This allows to study vertices of relatively high degree, e.g. close to the maximum degree (note that there are a.a.s. no vertices with degree exactly the expected maximum degree of $k_n = n^{\frac{1}{2\alpha}}$).}

%The main difference between Theorem \ref{thm:fixed_local_clustering_hyperbolic} and Theorem \ref{thm:local_clustering_hyperbolic} is that for fixed degrees $k$ we have a small correction term that depends on the value $k$.

>>>>

The aim of this paper is to study local clustering in hyperbolic random graphs. The first rigorous mathematical analysis of the local clustering coefficient was done by Gugelmann et al.~\cite{gugelmann2012random}. They show~\cite[Theorem 2.1]{gugelmann2012random} that the local clustering coefficient\footnote{Note that in~\cite{gugelmann2012random} this is called the global clustering coefficient. However, since this term is more often associated in the literature with the density of triangles compared to the number of paths of length 2, we use the term local clustering coefficient as is done, for instance, in~\cite{candellero2016clustering}.} is concentrated around its expectation which is shown to be $\bigT{1}$ as the number of vertices tends to infinity. In particular this result implies that the local clustering coefficient is asymptotically bounded away from zero. The convergence of this coefficient was not proven. In contrast, the global clustering coefficient was shown~\cite{candellero2016clustering} to converge in probability to some constant which can be explicitly stated as an integral expression, see Theorem 1.2 in \cite{candellero2016clustering}. The authors mention, however, that their analysis needs significant modification to be able to deal with convergence of the local clustering coefficient.

@@

A major part of this paper is dedicated to prove the convergence result in Theorem~\ref{thm:local_clustering_hyperbolic}, which is based on a sequence of subsequent results for the local clustering function in related random graph models and Theorem~\ref{thm:asymptotics_average_clustering_P}. The exact flow of the argument is explained in Section \ref{sec:proof_outline}. The proof for the convergence result in Theorem~\ref{thm:local_clustering_hyperbolic}, using these results, can be found in Section \ref{ssec:proof_main_results}. Here we also give the proof of the convergence result in Theorem~\ref{thm:clustering_coefficient_hyperbolic}. Another major part of this paper is the exact computation of the local clustering coefficient and function limit $c_\infty$ and $c_\infty(k)$, which involves careful computations of several involved integrals. This is done in Section~\ref{sec:asymptotics_average_clustering_ast_P}, where we also prove Theorem~\ref{thm:asymptotics_average_clustering_P}.

@@

\paragraph{Local clustering function for growing versus fixed degrees}

@@

%\paragraph{Transition in the scaling of local clustering}
%Note that Theorem \ref{thm:local_clustering_hyperbolic} implies that for $\frac{1}{2} < \alpha < \frac{3}{4}$ the local clustering $c_{\H,n}(k)$ decays as $k^{-(4\alpha - 2)}$ while for $\alpha > \frac{3}{4}$ it is $k^{-1}$.

%\paragraph{Universality of scaling transition}

%\paragraph{Geometric inhomogeneous random graphs}

### 2018-11-05 paper.tex
cleaned up proof of Theorem 1.5 and realigned formulas in section 4.5
Started merging Nikolaos results into paper
merged more of Nikoloas' computations into paper.
Section 4 done (plus some small changes to Section 1.5)

\newcommand{\eps}{\varepsilon}						%	varepsilon
\DeclareMathOperator{\dd}{\, \mathrm{d}\hspace{-1pt}}			%	straight d for integration

@@

\renewcommand{\H}{\ensuremath{\mathbb{H}}}			%	H used for the hyperbolic plane and hyperbolic random graph
\newcommand{\HP}{\ensuremath{\widetilde{\H}}}		%	Poisson hyperbolic random graph
\newcommand{\Dcal}[0]{\ensuremath{{\mathcal D}}}	%	Hyperbolic disc
\newcommand{\Rcal}[0]{\ensuremath{{\mathcal R}}}	%	Image of the hyperbolic disc under \Psi
\newcommand{\Pcal}[0]{\ensuremath{{\mathcal P}}}	%	Point process
\newcommand{\Po}[0]{\ensuremath{{\mathrm{Po}}}}		%	Poisson random variable

@@

@@

\newcommand{\hsym}[1]{h_{\H \bigtriangleup \Pcal}(#1)}
\newcommand{\EdgeDiff}{E_{\H \setminus \mathcal{P}}}
\newcommand{\expH}{\overline{n}_{\H}(k_n)}
\newcommand{\expP}{\overline{n}_{\Pcal}(k_n)}

@@

\PvdH{Below is a list of high and low level to do items. Please add comments using the commands below.

\noindent Comment commands:\\
Markus: \textbackslash MS\\
Nikolaos: \textbackslash NF\\
Pim: \textbackslash PvdH\\
Tobias: \textbackslash TM

}

\ToDo{High level list:

\begin{itemize}
\item Streamline notations and writing of some sections
\item Finish introduction
\item Complete bibliography list
\item Add additional explanation for the Hyperbolic model.
\item Add additional observations based on our main result
\end{itemize}
}

>>>>

\newcommand{\eps}{\varepsilon}						%	varepsilon
\DeclareMathOperator{\dd}{\mathrm{d}\hspace{-2pt}}			%	straight d for integration

@@

\renewcommand{\H}{\ensuremath{\mathbb{H}}}			%	H used for the hyperbolic plane and hyperbolic random graph
\newcommand{\HP}{\ensuremath{\widetilde{\H}}}		%	Poisson hyperbolic random graph
\newcommand{\Dcal}[0]{\ensuremath{{\mathcal D}}}	%	Hyperbolic disc
\newcommand{\Rcal}[0]{\ensuremath{{\mathcal R}}}	%	Image of the hyperbolic disc under \Psi
\newcommand{\Pcal}[0]{\ensuremath{{\mathcal P}}}	%	Point process
\newcommand{\Po}[0]{\ensuremath{{\mathrm{Po}}}}		%	Poisson random variable
\newcommand{\PPP}{\ensuremath{\Pcal_{\alpha, \nu}}}	%	Poisson Point Process

@@

\newcommand{\dom}[1]{\mathcal{D}_{#1}}

@@

\newcommand{\hsym}[1]{h_{\H \bigtriangleup \Pcal}(#1)}
\newcommand{\EdgeDiff}{E_{\H \setminus \mathcal{P}}}
\newcommand{\expH}{\overline{n}_{\HP}(k_n)}
\newcommand{\expP}{\overline{n}_{\Pcal}(k_n)}
\newcommand{\Pop}[2]{\mathrm{Po} (#2;#1)}
\newcommand{\Nsymdiff}{N_{\H \bigtriangleup \Pcal}(k_n)}

@@

### 2018-11-05 proof_overview.tex
Section 4 done (plus some small changes to Section 1.5)

The key result in this paper is Theorem~\ref{thm:local_clustering_hyperbolic}. In particular, it is the key ingredient in the proofs of Theorem~\ref{thm:clustering_coefficient_hyperbolic} and Theorem~\ref{thm:asymptotics_average_clustering_P}. Therefore, in this section we first give a detailed overview of the structure for the proof of Theorem~\ref{thm:local_clustering_hyperbolic}. The proof itself, based on intermediate propositions, can be found in Section~\ref{ssec:proof_main_results}. Here we also prove Theorem~\ref{thm:clustering_coefficient_hyperbolic}. The proof of Theorem~\ref{thm:asymptotics_average_clustering_P} can be found in Section~\ref{sec:asymptotics_average_clustering_ast_P}.

@@

Then, since $\Exp{\Prob{D_\Pcal = t}} = \bigT{1} t^{-(2\alpha + 1)}$
\begin{align*}
	&\hspace{-30pt}\sum_{t = 3}^{a_n} \Prob{D_\Pcal = t} \Exp{\left|c_{\H,n}^\ast(t) - c_{\Pcal}(t)\right|}\\
	&\le \bigT{1} 3^{-(2\alpha+1)} a_n \max_{3 \le t \le a_n} \Exp{\left|c_{\H,n}^\ast(t) - c_{\Pcal}(t)\right|}
	= \smallO{1},
\end{align*}
while $\Exp{c_\Pcal(t)} \le 1$ implies that
\begin{align*}
	&\hspace{-30pt}\frac{1}{n} \sum_{t = 3}^{a_n} \Exp{\left|N_{\H,n}(t) - n \Prob{D_\Pcal = t}\right|}\Exp{c_\Pcal(t)} \\
	&\le \sum_{t = 3}^{a_n} \Prob{D_\Pcal = t} \Exp{\left|\frac{N_{\H,n}(t)}{n \Prob{D_\Pcal = t}} - 1\right|} \\
	&\le \bigT{1} 3^{-(2\alpha+1)} a_n \max_{3 \le t \le a_n} 
		\Exp{\left|\frac{N_{\H,n}(t)}{n \Prob{D_\Pcal = t}} - 1\right|} = \smallO{1}.
\end{align*}
This finishes the proof.
\end{proof}

>>>>

The key result in this paper is Theorem~\ref{thm:local_clustering_hyperbolic}. In particular, it is the key ingredient in the proofs of Theorem~\ref{thm:clustering_coefficient_hyperbolic} and Theorem~\ref{thm:asymptotics_average_clustering_P}. Therefore, in this section we first give a detailed overview of the structure for the proof of Theorem~\ref{thm:local_clustering_hyperbolic}. The proof itself, based on intermediate propositions, can be found in Section~\ref{ssec:proof_main_results}. Here we also prove Theorem~\ref{thm:clustering_coefficient_hyperbolic}. The proof of Theorem~\ref{thm:asymptotics_average_clustering_P} can be found in Section~\ref{sec:asymptotics_average_clustering_ast_P} where we also derive the exact expressions for $c_\infty$ and $c_\infty(k)$.

@@

Then, since $\Exp{\Prob{D_\Pcal = t}} = \bigT{1} t^{-(2\alpha + 1)}$
\begin{align*}
	&\hspace{-30pt}\sum_{t = 3}^{a_n} \Prob{D_\Pcal = t} \Exp{\left|c_{\H,n}^\ast(t) - c_{\Pcal}(t)\right|}\\
	&\le \bigT{1} 3^{-(2\alpha+1)} a_n \max_{3 \le t \le a_n} \Exp{\left|c_{\H,n}^\ast(t) - c_{\Pcal}(t)\right|}
	= \smallO{1},
\end{align*}
while $\Exp{c_\Pcal(t)} \le 1$ implies that
\begin{align*}
	&\hspace{-30pt}\frac{1}{n} \sum_{t = 3}^{a_n} \Exp{\left|N_{\H,n}(t) - n \Prob{D_\Pcal = t}\right|}\Exp{c_\Pcal(t)} \\
	&\le \sum_{t = 3}^{a_n} \Prob{D_\Pcal = t} \Exp{\left|\frac{N_{\H,n}(t)}{n \Prob{D_\Pcal = t}} - 1\right|} \\
	&\le \bigT{1} 3^{-(2\alpha+1)} a_n \max_{3 \le t \le a_n} 
		\Exp{\left|\frac{N_{\H,n}(t)}{n \Prob{D_\Pcal = t}} - 1\right|} = \smallO{1}.
\end{align*}
This finishes the first part of the proof. The second part, which derives the exact expression for $c_\Pcal$ can be found in [??] \PvdH{Add reference to section.}
\end{proof}
